Meta-analysis

### 2023-03-27 analysis/meta-analysis.Rmd

```{r}
pool.groups = function(n1, n2, m1, m2, sd1, sd2) {

@@

## Step 1:
Combined different languages within the same study using the pooling formula below on the raw metrics:

Zheng & Samuel (2018) Korean and English

@@

```{r}
combine_lang <- function(data, study, measure, measure_overall, lang1, lang2, Tonal = T) {
  dat <- data %>%
    filter(Study == study) %>%
    filter(Measure == measure) %>%
    select(Tonal_N, Tonal.Mean, Tonal.SD)

@@

## Step 2:
calculated cohen's d, hedge's g, se of cohen'd and hedge's g

```{r}
sd_pooled <- function(n1, n2, s1, s2) {
  sd = sqrt(
    ((n1-1)*s1^2 + (n2-1)*s2^2)/
      ((n1-1) + (n2-1))
    )
  return(sd)
}

@@

meta_grouped$cohen_d <- ifelse(is.na(meta_grouped$cohen_d), meta_grouped$Cohen.s.D, meta_grouped$cohen_d) # add Alexander2008 to cohen's d list

@@

esc_mean_sd(grp1m = 0.8060976, grp2m = 0.6600000,
            grp1sd = 0.18554480, grp2sd = 0.2000000,
            grp1n = 41, grp2n = 26) #checking that formula is coded correctly
```

@@

See Page 228 in the intro to meta-analysis book for equations

For now, will assume the correlation is 1 between these subtasks

```{r}
#var_combined <- function(vector) {
#  n <- length(vector)
#  V1 <- vector[1]
#  V2 <- vector[2]
#  V3 <- ifelse(n >2, vector[3], 0)
#  return(((1/n)**2)*(V1 + V2 + V3 + 2*sqrt(V1*V2) + 2*sqrt(V1*V3) +2*sqrt(V2*V3)))
#}

sum_calc <- function(vec){
  sum = 0
  for (i in c(1:length(vec))){
  for(j in c(1:length(vec))){
    if(j == i){
      sum = sum
    }
    else{
      sum = sum + sqrt(vec[i]*vec[j])
    }
    j = j+1
  }
  i = i + 1
  }
  return(sum)
}

@@

```{r}
# Chen 2016: Accuracy and raw score not 100% consistent (maybe due to excluding trials). Will calculate using accuracy bc it takes into account trial-excluding
meta_processed <- meta_processed %>%
  filter(!(Study == "Chen et al. (2016)" & Measure == "Melody subtest"))
```

@@

```{r}
# Removing rhythm tests, meter, memory, speed, and non-traditional (nonspeech pitch contour identification) tests for now
meta_filtered <- meta_processed %>%
  #filter(Measure.Classification != "Rhythm") %>%
  #filter(Measure != "Rhythm test") %>%
  filter(!Measure %in% c("Pitch processing speed", "Pitch memory", "Nonspeech pitch contour identification", "Meter", "Memory"))

meta_filtered <- meta_filtered %>%
  mutate(hedge_g = case_when(
    Flip == T ~ -1*hedge_g,
    TRUE~hedge_g))
```

@@

```{r}
#meta_filtered <- meta_filtered %>%
#  filter(Study != "Chang et al. (2016)") # removing this study as an outlier

meta_processed_final <- meta_filtered %>%
  filter(Study != "Zhang et al. (2020)") # removing study comparing Japanese and Mandarin
```

@@

```{r}
meta_processed_final %>%
  count(Measure.Classification)
```

@@

melodic.gen.random  <- metagen(TE = hedge_g,
                 seTE = se_g, #
                 studlab = Study_Measure,
                 data = meta_melodic,
                 sm = "SMD",
                 fixed = FALSE,
                 random = TRUE,
                 method.tau = "REML",
                 hakn = TRUE,
                 title = "tonelang")

@@

```{r}
# Adding in language information 
nonCM_tonal <- as.data.frame(cbind(Study = c("Bradley (2016)",
                "Swaminathan et al (2021)",
                "Pfordresher \\& Brown (2009)",
                "Peretz et al. (2011)",
                "Ngo et al. (2016)",
                "Stevens et al. (2013)",
                "Hove et al. (2010)"),
      Non_CM_tonal = c("Yoruba, 15",
                       "Vietnamese, 3",
                       "Vietnamese, 6",
                       "Vietnamese, 4",
                       "Vietnamese, 8",
                       "Thai, 24",
                       "Hmong, 14"),
      Tonal_language = c("Mandarin or Yoruba",
                         "Mandarin or Cantonese (mostly) vs English (mostly)",
                         "Mandarin, Cantonese, or Vietnamese",
                         "Mandarin, Cantonese, or Vietnamese",
                         "Vietnamese",
                         "Thai",
                         "Chinese or Hmong"))
)

meta_processed_lang <- full_join(meta_processed_final, nonCM_tonal) %>%
  mutate(Non_CM_tonal = replace_na(Non_CM_tonal, "No"))
```

```{r}
#save(list = c("melodic.gen.random", "pitch.gen.random", "rhythm.gen.random", "study_num", "study_total_num"), file = here("results", "meta_analyses.RData"))
```

>>>>

```{r}
# function that pools two different groups
pool.groups = function(n1, n2, m1, m2, sd1, sd2) {

@@

## Step 1:

Combine different languages within the same study and language type (tonal or non-tonal) using the pooling formula below on the raw metrics

@@

```{r}
# function that pools language groups
combine_lang <- function(data, study, measure, measure_overall, lang1, lang2, Tonal = T) {
  dat <- data %>%
    filter(Study == study) %>%
    filter(Measure == measure) %>%
    select(Tonal_N, Tonal.Mean, Tonal.SD)

@@

## Step 2:

calculate cohen's d, hedge's g and their SE

```{r}
sd_pooled <- function(n1, n2, s1, s2) { # function that calculates the pooled sd of tonal and non-tonal for each study/entry
  sd = sqrt(
    ((n1-1)*s1^2 + (n2-1)*s2^2)/
      ((n1-1) + (n2-1))
    )
  return(sd)
}

@@

meta_grouped$cohen_d <- ifelse(is.na(meta_grouped$cohen_d), meta_grouped$Cohen.s.D, meta_grouped$cohen_d) # add Alexander2008 to cohen's d list; this study reported cohen's d directly

@@

esc_mean_sd(grp1m = 0.8060976, grp2m = 0.6600000,
            grp1sd = 0.18554480, grp2sd = 0.2000000,
            grp1n = 41, grp2n = 26) # checking that formula is coded correctly using Melodic comparison from Bradley (2016) as an example
```

@@

Next, we combine subtasks effect sizes into a total effect size for studies that broke down results at subtask levels (e.g., 5 vs. 10 cents difference)

Calculation method comes from Page 228 of Borenstein, M., Hedges, L. V., Higgins, J. P. T., & Rothstein, H. R. (2009). Introduction to meta-analysis. John Wiley & Sons

The correlation is assumed to be 1 between these subtasks

```{r}
sum_calc <- function(vec){
  sum = 0
  for (i in c(1:length(vec))){
  for(j in c(1:length(vec))){
    if(j == i){
      sum = sum
    }
    else{
      sum = sum + sqrt(vec[i]*vec[j])
    }
    j = j+1
  }
  i = i + 1
  }
  return(sum)
}

@@

```{r}
# Chen 2016: Reported both accuracy and raw score. Accuracy and raw score not 100% consistent (maybe due to excluding trials). Will calculate using accuracy since it takes into account trial-excluding
meta_processed <- meta_processed %>%
  filter(!(Study == "Chen et al. (2016)" & Measure == "Melody subtest"))
```

@@

```{r}
# Removing rhythm tests, meter, memory, speed, and non-traditional (nonspeech pitch contour identification) tests for now
meta_filtered <- meta_processed %>%
  filter(!Measure %in% c("Pitch processing speed", "Pitch memory", "Nonspeech pitch contour identification", "Meter", "Memory"))

meta_filtered <- meta_filtered %>%
  mutate(hedge_g = case_when(
    Flip == T ~ -1*hedge_g, # tasks with flip == T are those that are reverse coded (high value == low performance)
    TRUE~hedge_g))
```

@@

```{r}
meta_processed_final <- meta_filtered %>%
  filter(Study != "Zhang et al. (2020)") # removing study comparing Japanese and Mandarin
```

@@

```{r}
meta_processed_final %>%
  count(Measure.Classification) # number of tasks per category
```

@@

melodic.gen.random  <- metagen(TE = hedge_g,
                 seTE = se_g, 
                 studlab = Study_Measure,
                 data = meta_melodic,
                 sm = "SMD",
                 fixed = FALSE,
                 random = TRUE,
                 method.tau = "REML",
                 hakn = TRUE,
                 title = "tonelang")

@@

```{r}
# save(list = c("melodic.gen.random", "pitch.gen.random", "rhythm.gen.random", "study_num", "study_total_num"), file = here("results", "meta_analyses.RData"))
# uncomment to save 
```
